---
title: "Model Checking"
author: "Azoacha Forcheh, 20558994"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
market_origin <- read.csv("market_index_clean.csv")
first <- market_origin[1,]
market <- market_origin[2:3104,]
market$AAPL2 <- market_origin[,"AAPL"][1:3103]
market$Return <- (market$AAPL - market$AAPL2)/market$AAPL2
```

```{r}
library(lubridate)
market$Date <- market_origin$Date[2:3104]
date <- market$Date
first <- date[1]
day <- c()
for(i in 1:length(date)) {
  diff = round(difftime(date[i], first))
  day = c(day, diff)
}
market$Date <- day
fit <- lm(Return~., data = market)
summary(fit)
```

```{r}
nullmodel <- lm(Return~1, data=market)
fullmodel <- lm(Return~., data=market)
step(nullmodel, scope=list(upper=fullmodel), direction="both")
```


```{r}
# Check for multicollinearity
# Date , SPX , BNDGLB , EEM, and AAPL2 have VIF >> 10 (consider dropping)
library(mctest)
imcdiag(market[ , !(names(market) %in% c('Return'))], market$Return)
```

```{r fig.width=10, fig.height=8}
test_model <- lm(Return ~ VIX + AAPL2 + AAPL + SPGSCITR, data = market)
par(mfrow=c(2,2))
plot(test_model)
summary(test_model)
#plot(market$Return, residuals(test_model), xlab = "Return ratio", ylab = "Residuals")
#abline(a = 0, b= 0, col = "red")
```

```{r fig.width=10, fig.height=8}
library(car)
crPlots(m=test_model, main="Partial Residual Plots", 
        ylab="Partial Residual", smooth=T)
```

## Notes:
## No real influential cases in Residuals/Leverage plot; can barely see Cook's distance lines. No influential outliers.
## Normal Q-Q is light-tailed, and hence the consequences of non-normality is not serious and can be ignored.
## None of the covariates are significantly deviating from the straight line. However, residuals plot looks like a y = -x^3 graph,
##  and this curvature indicates the need for a Boxcox transformation.

```{r}
# Autocorrelation tests
library(lmtest)
acf(residuals(test_model), main="Residuals", lag.max=10)
dwtest(test_model, alternative=c("two.sided"), data=market)
```

# Autocorrelation in the response has been explained away - plot indicates no autocorrelation, and 
#   Durbin-Watson test shows we have evidence to accept null hypothesis at 95% significance level.



