---
title: \bf{Project B Report}
output: 
  pdf_document: 
    latex_engine: xelatex
---

\begin{section}{Summary}
The objective of this report is to model the Apple Stock return rate, defined as $ret.AAPL_i = \frac{AAPL_i - AAPL_{i-1}}{AAPL_{i-1}}$, based on key metrics such as volatility, $S\&P$ 500 Index, and etc. 

To model the return rate, our group has surveyed a wide range of models including an autoregressive model, linear models from automated model selection, as well as transformed linear models. 

From the autoregressive model, we noticed the heteroscedascity caused by autocorrelation, which led us to determine that directly modelling the stock price is not the most ideal way to model the return rate. We also found discrepancy between the adjusted R-Squared value and the residual plots in the backfitted model, which will be further discussed later in this report. Through our investigation, our group has concluded that there is no truly ideal linear model for the Apple Stock return rate. Nevertheless, we were able to identify the inadequate models and gain insights on why those models did not work.
\end{section}

\begin{section}{Model Selection}
The first model to be considered is the autoregressive model, which directly models the stock price of day $i$ using the price from day $i-1$ and has the following expression: $$ AAPL_i = \beta_0 + \beta_1 AAPL_{i-1}$$

The residual plot from the fitted model has a clearly heteroscedastic property, since the residuals take on a fan shape.

\begin{center}\includegraphics[width=90mm]{autoregressive_residual_vs_fitted.png}\end{center}

We tried to resolve the heteroscedascity issue by implementing a Weighted Least Squares autoregressive model but the results remained the same (see Appendix for full code and plot comparison). After some research, it turns out that this is an irreducible property of the autoregressive model because the data is serially correlated. We realized that we needed something better than directly modelling the price. Hence we decided to directly model the return which, to re-iterate, is defined as $ret.AAPL_i = \frac{AAPL_i - AAPL_{i-1}}{AAPL_{i-1}}$.

Correlation between the coefficients was also expected from the data. Since it is a time-series financial data set, we knew and verified that time is correlated to variables such as the Apple Stock Price and the $S \& P$ 500 Index. The detailed matrix can be found in the Appendix section. We decided that this wasn't too much of an issue since it is an inevitable part of any time-series data.

Moving on from the autoregressive model, we then considered automated model selecting using forward selection, backward elemination, and stepwise selection, from which we obtained the following models (full code in Appendix):


\textbf{Forward selection: } $Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3AAPL + \beta_4SPGSCITR$

\textbf{Backward elimination: } $Return = \beta_0 + \beta_1AAPL + \beta_2VIX + \beta_3SPGSCITR + \beta_4AAPL2$

\textbf{Stepwise selection: } $Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3AAPL + \beta_4SPGSCITR$


Interestingly, all selection methods returned the same model. Before we proceeded any further, we consulted with Prof. Zeng. After receiving helpful advice from the professor, we realized that since $Return$ is calculated as a function of $AAPL$ and $AAPL2$, the two variables cannot be included in the same model. We then adjusted the full model by excluding the variable $AAPL$, hence only keeping the previous-day price, and obtained the results below:


\textbf{Forward selection: } $ Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3SPGSCITR + \beta_4EEM$

\textbf{Backward elimination: } $ Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3SPGSCITR + \beta_4EEM$

\textbf{Stepwise selection: } $ Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3SPGSCITR + \beta_4EEM$

We decided to further analyze this model in the model diagnostics section.

Additionally, we considered how the distribution of the independent variables could affect the model fit. Since both $VIX$ and $SPGSCITR$ had heavily skewed histograms, we decided to log-transform the variables to make the distributions resemble a normal curve and obtained the following model:
\textbf{Transformed model: } $ Return = \beta_0 + \beta_1(log(VIX)^{-2}) + \beta_2AAPL2 + \beta_3(log(SPGSCITR)) + \beta_4EEM$. This will be the second model to be discussed in the model diagnostics section.

\end{section}

\begin{section}{Model Diagnostics}
\end{section}

\begin{section}{Discussion}

Following questions are copied from the given project instructions.

\begin{enumerate}
\item{}
\textit {What are the most important factors affecting Apple Stock returns?}

According to the results from Rcodes below, we can check the $p$-values for the significance of coefficients with the corresponding covariates. For the transformed model, $VIX$ term has the lowest $p$-value as shown in the summary, which means it would make largest change to the response if it drops out of the model; while $AAPL2$ term, which is the pervious day's price, and $SPGSCITR$ term's coefficients ealso have relatively small $p$-values. For our selected model before transformation, $VIX$ has the lowest p-value as well, and again $AAPL2$ and $SPGSCITR$ terms' coefficients have considerably small $p$-values too.  So based on our models, $VIX$ would be the most important factor, and $SPGSCITR$ and $AAPL2$ are considerably significant as well. 

\item{}
\textit {Does including information for several days ago have much of an impact on tomorrow’s return, if we account for today’s information?}

As shown in the R code in the Appendix, after attempting to instead of using previous day's price to calculate return, we take the average of the previous 5 days' stock prices and still fitting the same transformed model. Comparing to original transformed model, the performance of the model seems to be better in temrs of coefficient of determination and the AIC values, while there is a small increase in adjusted $R^2$ and a small decrease for AIC. So it seems that from our attempt, including information for several days ago would give a better model fitting the tomorrow's return, while we are still accounting for today's information.

\item{}
\textit {Does the model predict accurately?}

Based on our results on our k-fold cross validation with the comparsion of training and testing sets of data, the model performs generally well for prediction accuracy. Although in the summary of the model, the coefficient of determination is not ideally high, as we have discussed in class and the fact that $R^2$ can be affected by other factors (such as number of variables that are involved), so it would be acceptable to say that the model predicts the future value with reasonably accurate results. 

\item{}
\textit {Are there any coefficients with high $p$-values retained in the final model? If so, why?}

According our models, fortunately we don't have any coefficients with high $p$-values retained in the final model. All of the coefficients have descent $p$-values that are lower than 0.05.

\item{}
\textit {Are any of the regression assumptions of the final model violated? If so, which ones?}

This question has been mentioned along with the Model Diagnostics part. All of the regression assumptions seems to be satisfied by our final model, instead of there is a little bit bias on the independence of residuals.

\item{}
\textit {What are the possible deficiencies of the final model? How the model can be improved?}

As discussed in previous analysis part, and also by looking at the residual plots and the summary of aborted models, the data has a tendency to be fitting a non-linear model. For example, there is a curved trend for the residual against fitted values plot, even after the chosen transformation. And notice that from the R results we can see that the adjusted and unadjusted coefficients of determination is very small and not close to 1 at all, which is not ideal as well. To improve the fittness of the model, it might make more sense to fit a non-linear model instead. Or adding more covariates to the model, but that would include the possiblity of overfitting the model.  

\item{}
\textit {Are there any outlying observations that might be appropriate to remove?}

By judging from the results of residual plots and leverage plots, it can be seen that there seems to have several outliers concentrated around the year 2008, which has the index 1370. The possible reason for the outlying results could be the finacial crisis during that time that would cause the drop in variates' values, but AAPL stock price didn't be affected that much, possible explanation could be the release of new series of iPhone which sort of stablize the change in stock price for Apple. Under such circumstance, there might exist possible reasons for the unexpected change, here it would be more reasonable to ignore the outlying results, as it is not really "unexpected" or it is abnormal. Therefore it would not be appropriate to remove.

\end{enumerate}
\end{section}

\newpage
\begin{section}{Appendix}
Below is a collection of R code used in this report.
\end{section}

### Pre-process the data
```{r}
market_origin <- read.csv("market_index_clean.csv")
first <- market_origin[1,]
market <- market_origin[2:3104,]
market$AAPL2 <- market_origin[,"AAPL"][1:3103]
market$Return <- (market$AAPL - market$AAPL2)/market$AAPL2
```

### Full Model
```{r}
library(lubridate)
market$Date <- market_origin$Date[2:3104]
date <- market$Date
first <- date[1]
day <- c()
for(i in 1:length(date)) {
  diff = round(difftime(date[i], first))
  day = c(day, diff)
}
market$Date <- day
full_model <- lm(Return~., data = market)
summary(full_model)
```

### Null Model
```{r}
null_model <- lm(Return~1, data = market)
summary(null_model)
```

### Autoregressive Model
```{r}
autoregressive_model <- lm(AAPL~AAPL2, data = market)
plot(fitted(autoregressive_model),residuals(autoregressive_model), 
     main = "Residual vs. Fitted", xlab = "Fitted Values", ylab = "Residuals")
```

### WLS Autoregressive Model
```{r}
wts <-1/(fitted(lm(abs(residuals(autoregressive_model))~fitted(autoregressive_model)))^2)
wls <- lm(market$AAPL~market$AAPL2, weights = wts)
par(mfrow=c(1,2))
plot(fitted(autoregressive_model), autoregressive_model$residuals, 
     main="unweighted", xlab = "fitted", ylab = "residuals")
plot(fitted(wls), wls$residuals, main="weighted", xlab = "fitted", ylab = "residuals")
```

There is no apparent difference between the weighted least squares and the un-weighted autoregressive models.

### Correlation Matrix
```{r}
cor(market)
```

### Forward Selection
```{r}
step(null_model, scope = list(upper=full_model), direction="forward")
summary(lm(formula = Return ~ VIX + AAPL2 + AAPL + SPGSCITR, data = market))
```

### Backward Elimination
```{r}
step(full_model, scope = list(lower=null_model), direction="backward")
summary(lm(formula = Return ~ AAPL + VIX + SPGSCITR + AAPL2, data = market))
```

### Stepwise Selection
```{r}
step(null_model, scope = list(upper=full_model), direction="both")
```


### Forward Selection without AAPL
```{r}
full_model <- lm(Return~SPX+VIX+SPGSCITR+BNDGLB+EEM+AAPL2, data = market)
step(null_model, scope = list(upper=full_model), direction="forward")
```

### Backward Elimination without AAPL
```{r}
step(full_model, scope = list(lower=null_model), direction="backward")
```

### Stepwise Selection without AAPL
```{r}
step(null_model, scope = list(upper=full_model), direction="both")
```

### Log-transformed model from stepwise selection results
```{r}
## note that the VIX needed to be raised to the -2, after the log transformation
##  in order to obtain a distribution close to the Gaussian curve
VIXtr <- log(market$VIX)^-2
hist(VIXtr)
SPGSCITRtr <- log(market$SPGSCITR)
hist(SPGSCITRtr)
transformed_model <- lm(market$Return~VIXtr+SPGSCITRtr+market$EEM+market$AAPL2)
summary(transformed_model)
```

### The impact of including more data from previous days
```{r}
## suppose we include previous 5 days data information

## attempt to using the average value of previous 5 days AAPL
AAPL5 <- c()
for (i in 1:(nrow(market_origin) - 4)){
  set <- c(market_origin$AAPL[i], market_origin$AAPL[i+1], market_origin$AAPL[i+2], market_origin$AAPL[i+3], market_origin$AAPL[i+4])
  AAPL5 <- c(AAPL5, mean(set))
}
new_AAPL = market$AAPL[4:length(market$AAPL)]
new_Return <- (new_AAPL - AAPL5)/new_AAPL
new_VIXtr = VIXtr[4:length(VIXtr)]
new_SPGSCITRtr = SPGSCITRtr[4:length(SPGSCITRtr)]
new_EEM = market$EEM[4:length(market$EEM)]
new_transformed_model <- lm(new_Return~new_VIXtr+new_SPGSCITRtr+new_EEM+AAPL5)
summary(new_transformed_model)
AIC(new_transformed_model)
AIC(transformed_model)
```



