---
title: \bf{Project B Report}
output: 
  pdf_document: 
    latex_engine: xelatex
---
\begin{section}{Delete this section before submitting}
These are the three questions specific to Project B. We need to find appropriate sections to answer these questions (and make it clear that we're answering them so that the TA gives us marks)
\begin{enumerate}
\item{}
What are the most important factors affecting Apple Stock returns?
\item{}
Does including information for several days ago have much of an impact on tomorrow’s
return, if we account for today’s information?
\item{}
Does the model predict accurately?
\end{enumerate}
\end{section}

\begin{section}{Summary}
The objective of this report is to model the Apple Stock return rate, defined as $ret.AAPL_i = \frac{AAPL_i - AAPL_{i-1}}{AAPL_{i-1}}$, based on key metrics such as volatility, $S\&P$ 500 Index, and etc. 

To model the return rate, our group has surveyed a wide range of models including an autoregressive model, linear models from automated model selection, as well as transformed linear models. 

From the autoregressive model, we noticed the heteroscedascity caused by autocorrelation, which led us to determine that directly modelling the stock price is not the most ideal way to model the return rate. We also found discrepancy between the adjusted R-Squared value and the residual plots in the backfitted model, which will be further discussed later in this report. Through our investigation, our group has concluded that there is no truly ideal linear model for the Apple Stock return rate. Nevertheless, we were able to identify the inadequate models and gain insights on why those models did not work.
\end{section}

\begin{section}{Model Selection}
The first model to be considered is the autoregressive model, which directly models the stock price of day $i$ using the price from day $i-1$ and has the following expression: $$ AAPL_i = \beta_0 + \beta_1 AAPL_{i-1}$$

The residual plot from the fitted model has a clearly heteroscedastic property, since the residuals take on a fan shape.

\begin{center}\includegraphics[width=90mm]{autoregressive_residual_vs_fitted.png}\end{center}

We tried to resolve the heteroscedascity issue by implementing a Weighted Least Squares autoregressive model but the results remained the same (see Appendix for full code and plot comparison). After some research, it turns out that this is an irreducible property of the autoregressive model because the data is serially correlated. We realized that we needed something better than directly modelling the price. Hence we decided to directly model the return which, to re-iterate, is defined as $ret.AAPL_i = \frac{AAPL_i - AAPL_{i-1}}{AAPL_{i-1}}$.

Correlation between the coefficients was also expected from the data. Since it is a time-series financial data set, we knew and verified that time is correlated to variables such as the Apple Stock Price and the $S \& P$ 500 Index. The detailed matrix can be found in the Appendix section. We decided that this wasn't too much of an issue since it is an inevitable part of any time-series data.

Moving on from the autoregressive model, we then considered automated model selecting using forward selection, backward elemination, and stepwise selection, from which we obtained the following models (full code in Appendix):


\textbf{Forward selection: } $Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3AAPL + \beta_4SPGSCITR$

\textbf{Backward elimination: } $Return = \beta_0 + \beta_1AAPL + \beta_2VIX + \beta_3SPGSCITR + \beta_4AAPL2$

\textbf{Stepwise selection: } $Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3AAPL + \beta_4SPGSCITR$


Interestingly, all selection methods returned the same model. Before we proceeded any further, we consulted with Prof. Zeng. After receiving helpful advice from the professor, we realized that since $Return$ is calculated as a function of $AAPL$ and $AAPL2$, the two variables cannot be included in the same model. We then adjusted the full model by excluding the variable $AAPL$, hence only keeping the previous-day price, and obtained the results below:


\textbf{Forward selection: } $ Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3SPGSCITR + \beta_4EEM$

\textbf{Backward elimination: } $ Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3SPGSCITR + \beta_4EEM$

\textbf{Stepwise selection: } $ Return = \beta_0 + \beta_1VIX + \beta_2AAPL2 + \beta_3SPGSCITR + \beta_4EEM$

We decided to further analyze this model in the model diagnostics section.

Additionally, we considered how the distribution of the independent variables could affect the model fit. Since both $VIX$ and $SPGSCITR$ had heavily skewed histograms, we decided to log-transform the variables to make the distributions resemble a normal curve and obtained the following model:
\textbf{Transformed model: } $ Return = \beta_0 + \beta_1(log(VIX)^{-2}) + \beta_2AAPL2 + \beta_3(log(SPGSCITR)) + \beta_4EEM$. This will be the second model to be discussed in the model diagnostics section.

\end{section}

\begin{section}{Model Diagnostics}
\end{section}

\begin{section}{Discussion}
\end{section}

\newpage
\begin{section}{Appendix}
Below is a collection of R code used in this report.
\end{section}

### Pre-process the data
```{r}
market_origin <- read.csv("market_index_clean.csv")
first <- market_origin[1,]
market <- market_origin[2:3104,]
market$AAPL2 <- market_origin[,"AAPL"][1:3103]
market$Return <- (market$AAPL - market$AAPL2)/market$AAPL2
```

### Full Model
```{r}
library(lubridate)
market$Date <- market_origin$Date[2:3104]
date <- market$Date
first <- date[1]
day <- c()
for(i in 1:length(date)) {
  diff = round(difftime(date[i], first))
  day = c(day, diff)
}
market$Date <- day
full_model <- lm(Return~., data = market)
summary(full_model)
```

### Null Model
```{r}
null_model <- lm(Return~1, data = market)
summary(null_model)
```

### Autoregressive Model
```{r}
autoregressive_model <- lm(AAPL~AAPL2, data = market)
plot(fitted(autoregressive_model),residuals(autoregressive_model), 
     main = "Residual vs. Fitted", xlab = "Fitted Values", ylab = "Residuals")
```

### WLS Autoregressive Model
```{r}
wts <-1/(fitted(lm(abs(residuals(autoregressive_model))~fitted(autoregressive_model)))^2)
wls <- lm(market$AAPL~market$AAPL2, weights = wts)
par(mfrow=c(1,2))
plot(fitted(autoregressive_model), autoregressive_model$residuals, 
     main="unweighted", xlab = "fitted", ylab = "residuals")
plot(fitted(wls), wls$residuals, main="weighted", xlab = "fitted", ylab = "residuals")
```

There is no apparent difference between the weighted least squares and the un-weighted autoregressive models.

### Correlation Matrix
```{r}
cor(market)
```

### Forward Selection
```{r}
step(null_model, scope = list(upper=full_model), direction="forward")
summary(lm(formula = Return ~ VIX + AAPL2 + AAPL + SPGSCITR, data = market))
```

### Backward Elimination
```{r}
step(full_model, scope = list(lower=null_model), direction="backward")
summary(lm(formula = Return ~ AAPL + VIX + SPGSCITR + AAPL2, data = market))
```

### Stepwise Selection
```{r}
step(null_model, scope = list(upper=full_model), direction="both")
```


### Forward Selection without AAPL
```{r}
full_model <- lm(Return~SPX+VIX+SPGSCITR+BNDGLB+EEM+AAPL2, data = market)
step(null_model, scope = list(upper=full_model), direction="forward")
```

### Backward Elimination without AAPL
```{r}
step(full_model, scope = list(lower=null_model), direction="backward")
```

### Stepwise Selection without AAPL
```{r}
step(null_model, scope = list(upper=full_model), direction="both")
```

### Log-transformed model from stepwise selection results
```{r}
## note that the VIX needed to be raised to the -2, after the log transformation
##  in order to obtain a distribution close to the Gaussian curve
VIXtr <- log(market$VIX)^-2
hist(VIXtr)
SPGSCITRtr <- log(market$SPGSCITR)
hist(SPGSCITRtr)
transformed_model <- lm(market$Return~VIXtr+SPGSCITRtr+market$EEM+market$AAPL2)
summary(transformed_model)
```