---
title: "Testing Updated Models"
author: "Azoacha Forcheh, 20558994"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
market_origin <- read.csv("market_index_clean.csv")
first <- market_origin[1,]
market <- market_origin[2:3104,]
market$AAPL2 <- market_origin[,"AAPL"][1:3103]
market$Return <- (market$AAPL - market$AAPL2)/market$AAPL2
market$VIX2 = market_origin[,"VIX"][1:3103]
market$Return = market$Return/sqrt(market$VIX2)
```

```{r}
library(lubridate)
market$Date <- market_origin$Date[2:3104]
date <- market$Date
first <- date[1]
day <- c()
for(i in 1:length(date)) {
  diff = round(difftime(date[i], first))
  day = c(day, diff)
}
market$Date <- day
```

```{r fig.width=10, fig.height=8}
fullmodel <- lm(Return~., data = market)
summary(fullmodel)
par(mfrow=c(2,2))
plot(fullmodel)
```


```{r}
# Helper function to avoid NA's when applying power 
#   to vector with negative values
exponent <- function(a, pow) (abs(a)^pow)*sign(a)
```

## BoxCox transformation to deal with nonlinearity
## Using 1/3 due to shape of residual-fitted plot
```{r fig.width=10, fig.height=8}
market.tr = market
market.tr$lReturn = market$Return^(2/3)
market.tr$lAAPL = market$AAPL^(1/3)
market.tr$lAAPL2 = market$AAPL2^(1/3)
market.tr$VIX = market$VIX^(-1)
market.tr$VIX2 = market$VIX2^(-1)

#market.tr$Return = NULL
#market.tr$AAPL = NULL
market.tr$AAPL2 = NULL
market.tr$VIX = NULL
market.tr$VIX2 = NULL

fullmodel2 = lm(Return~., data = market.tr)
par(mfrow=c(2,2))
plot(fullmodel2)
summary(fullmodel2)
```

## This full model gives a higher adjusted R^2 than the full model on the untransformed data.

## Backward elimination on transformed dataframe
```{r}
drop1(fullmodel2, test="F")
drop1(update(fullmodel2, .~.-VIX), test="F")
drop1(update(fullmodel2, .~.-VIX-SPGSCITR), test="F")
drop1(update(fullmodel2, .~.-VIX-SPGSCITR-EEM), test="F")
drop1(update(fullmodel2, .~.-VIX-SPGSCITR-EEM-Date-BNDGLB-SPX), test="F")
```

## Based on this, we choose lReturn ~ lAAPL + lAAPL2

## Model selection with step for comparison
```{r}
nullmodel2 = lm(lReturn~1, data=market.tr)
step(nullmodel2, scope=list(upper=fullmodel2), direction="both")
```

## Different result: we choose lReturn ~ VIX + SPX + SPGSCITR, which gives significantly worse results

```{r fig.width=10, fig.height=8}
test_model2 = lm(lReturn ~ lAAPL + lAAPL2, data=market.tr)
par(mfrow=c(2,2))
plot(test_model2)
summary(test_model2)
```


```{r}
# Autocorrelation tests
library(lmtest)
acf(residuals(test_model2), main="Residuals", lag.max=10)
dwtest(test_model2, alternative=c("two.sided"), data=market.tr)
```

## Autocorrelation in the response has still been explained away - plot indicates no autocorrelation, and 
##   Durbin-Watson test shows we have evidence to accept null hypothesis at 95% significance level.


```{r fig.width=10, fig.height=8}
market.tr1 = market
market.tr1$lReturn = exponent(market$Return,3)
market.tr1$lAAPL = exponent(market$AAPL,1/3)
market.tr1$lAAPL2 = exponent(market$AAPL2,-1/3)

market.tr1$Return = NULL
#market.tr1$AAPL = NULL
#market.tr1$AAPL2 = NULL

fullmodel3 = lm(lReturn~., data = market.tr1)
par(mfrow=c(2,2))
plot(fullmodel3)
summary(fullmodel3)
```


```{r}
drop1(fullmodel3, test="F")
drop1(update(fullmodel3, .~.-EEM), test="F")
drop1(update(fullmodel3, .~.-EEM-Date), test="F")
drop1(update(fullmodel3, .~.-EEM-Date-SPX), test="F")
drop1(update(fullmodel3, .~.-EEM-Date-SPX-VIX), test="F")
drop1(update(fullmodel3, .~.-EEM-Date-SPX-VIX-SPGSCITR), test="F")
```

## Results: lReturn ~ BNDGLB + lAAPL + lAAPL2

```{r fig.width=10, fig.height=8}
test_model3 = lm(lReturn ~ BNDGLB + lAAPL + lAAPL2, data=market.tr1)
par(mfrow=c(2,2))
plot(test_model3)
summary(test_model3)
```

